Substitution rate responses to changes in effective population size.

### 2020-05-09 manuscript/supp-mat.tex
Added Figures and updated manuscript

For a given genotype sequence, define phenotype $0 \leq \x \leq 1$ as the current proportion of sites in the unstable state.
After a mutation, given that only one site can change at a time, the absolute change of $\x$ is either $0$ or $\dx=\sfrac{1}{\Nsite}$.
Define $\rho_{\x}(\dx)$ as the probability to get a change of phenotype equal to $\dx$, if the current phenotype is $\x$:\\

@@

\dx &\text{ with probability } \rho_{\x}(\dx) = 1-x, \\
0 &\text{ with probability } \rho_{\x}(0) = \x \left[1 - \frac{1}{\Nstate - 1}\right], \\
-\dx &\text{ with probability } \rho_{\x}(-\dx) = \frac{\x}{\Nstate - 1}.\\

@@

$s(\x, \dx)$ is the selection coefficient of an effect $\dx$ if the current phenotype is $\x$:

@@

s(\x, \dx) &= \frac{f(\x + \dx) - f(\x)}{f( \x)}, \\
 &\simeq \frac{1}{f(\x)}\frac{ \partial f(\x) }{\partial \x} \dx, \\
 &\simeq \frac{ \partial \ln f(\x) }{\partial \x} \dx, \label{eq:s_from_fitness}

@@

where $f( \x)$ is the Wrightian fitness of phenotype $x$. And thus we also have:

@@

s(\x, -\dx) \simeq - s(\x, \dx) \text{ from eq. \ref{eq:s_from_fitness}} \label{eq:s_minus_deltax}, \\
\iff S(\x, -\dx) \simeq - S(\x, \dx), \label{eq:S_minus_deltax}

@@

where $S(\x\eq, \dx) = 4\Ne s (\x\eq, \dx)$ is the scaled selection coefficient.

@@

The probability of fixation of a mutation with effect $\dx$, for a resident phenotype $\x$ is :

@@

p_{\text{fix}}(\x, \dx) &= \frac{ 2 s(\x, \dx)}{1 - \e^{-4\Ne s(\x, \dx)}}, \\
 &= \frac{ 2 s(\x, \dx)}{1 - \e^{-S(\x, \dx)}}. \label{eq:pfix}

@@

p_{\text{fix}}(\x, 0) = \frac{ 1}{2 \Ne} \label{eq:pfix0}.

@@

\frac{p_{\text{fix}}(\x, \dx)}{p_{\text{fix}}(\x, 0)} & = \frac{ 2 \Ne 2 s(\x, \dx)}{1 - \e^{-S(\x, \dx)}} \text{ from eq. \ref{eq:pfix} and \ref{eq:pfix0}},\\
& = \frac{S(\x, \dx)}{1 - \e^{-S(\x, \dx)}} \label{eq:pfixratio}.

@@

At equilibrium phenotype $\x\eq$, the expected selection coefficient of mutation that reached fixation must be $0$:

@@

 0 = \E_{\dx} \left[ s(\x\eq, \dx) p_{\text{fix}}(\x\eq, \dx) \right], \\
\iff 0 = \frac{ 2 s(\x\eq, \dx)^2}{1 - \e^{-S(\x\eq, \dx)}}   \rho_{\x\eq}(\dx) + s(\x\eq, 0) \frac{ \rho_{\x\eq}(0)}{2\Ne} + \frac{ 2s(\x\eq, -\dx)^2}{1 - \e^{-S(\x\eq, -\dx)}} \rho_{\x\eq}(-\dx)\text{ from eq. \ref{eq:pfix} and \ref{eq:pfix0}}, \\
\Longrightarrow \frac{ 2s(\x\eq, \dx)^2}{1 - \e^{-S(\x\eq, \dx)}}   \rho_{\x\eq}(\dx) \simeq \frac{ - 2s(\x\eq, \dx)^2}{1 - \e^{S(\x\eq, \dx)}}   \rho_{\x\eq}(-\dx) \text{ from eq. \ref{eq:S_minus_deltax}}, \\
\iff \frac{ \rho_{\x\eq}(\dx)}{\rho_{\x\eq}(-\dx)} \simeq \e^{-S(\x\eq, \dx)} \frac{ \e^{-S(\x\eq, \dx)} - 1}{ \e^{-S(\x\eq, \dx)} \left( 1 - \e^{S(\x\eq, \dx)} \right)}, \\
\iff \ln \left( \frac{1 - \x\eq}{\x\eq} \right) + \ln (\Nstate-1) \simeq -S(\x\eq, \dx) \text{ from eq. \ref{eq:proba}} \label{eq:equilibrium}, \\
\iff \lambda_{K}(\x\eq) \simeq -S(\x\eq, \dx) \label{eq:equilibrium_lambda},

@@

where  $\lambda_{K}(\x\eq) = \ln \left( \frac{1 - \x\eq}{\x\eq} \right) + \ln (\Nstate-1)$.
\subsection{Substitution rate bias ($\bm{\dnds}$) at equilibrium}
The substitution rate of selected mutations relative to neutral mutations ($\dnds$) is by definition:

@@

\dnds & = \E_{\dx} \left[ \frac{p_{\text{fix}}(\x, \dx)}{p_{\text{fix}}(\x, 0)} \right], \\
 & = (1 - \x) \frac{ S(\x, \dx)}{1 - \e^{-S(\x, \dx)}} + \x \left(\frac{\Nstate - 2}{\Nstate - 1}\right) + \frac{\x}{\Nstate-1} \frac{ S(\x, -\dx)}{1 - \e^{-S(\x, -\dx)}} \text{ from eq. \ref{eq:proba}, \ref{eq:pfix} and \ref{eq:pfix0}}, \\
 & = (1 - \x) \frac{ S(\x, \dx)}{1 - \e^{-S(\x, \dx)}} - \frac{\x}{\Nstate-1}  \frac{ S(\x, \dx)}{1 - \e^{S(\x, \dx)}} +  \x \left(\frac{\Nstate - 2}{\Nstate - 1}\right) \text{ from eq. \ref{eq:S_minus_deltax}}.

@@

$\dnds\eq$ at equilibrium is then determined by the phenotype at equilibrium $\x\eq$:

@@

\dnds\eq &= (1 - \x\eq) \frac{ S(\x\eq, \dx)}{1 - \e^{-S(\x\eq, \dx)}} - \frac{\x\eq}{\Nstate-1} \frac{ S(\x\eq, \dx)}{1 - \e^{S(\x\eq, \dx)}} + \x\eq \left(\frac{\Nstate - 2}{\Nstate - 1}\right) , \\
 &= \x\eq \left[ \frac{2 (\x\eq-1)  \lambda_{K}(\x\eq) }{\Nstate (\x\eq-1)+1} + \frac{\Nstate - 2}{\Nstate - 1}\right] \text{ from eq. \ref{eq:equilibrium}}.

@@

Moreover, given that the number of state if large enough $\Nstate \gg 1$, the equilibrium $\dnds$ can be approximated as:

@@

\dnds\eq  & = \x\eq \left[ \frac{2 (\x\eq-1)  \lambda_{K}(\x\eq) }{\Nstate (\x\eq-1)+1} + \frac{\Nstate - 2}{\Nstate - 1}\right], \\
& \simeq \x\eq

@@

And the derivative of $\dnds\eq$ w.r.t to $\x\eq$ is: 

@@

\frac{\der \dnds\eq}{\der \x\eq} = 2 \left[ \frac{\Nstate (\x\eq- 1) + 1+\left[\Nstate (\x\eq-1)^2+2 \x\eq-1\right] \lambda_{K}(\x\eq) }{(\Nstate (\x\eq-1)+1)^2}\right] + \frac{\Nstate - 2}{\Nstate - 1} \label{eq:dw_dx}.

@@

Moreover, given that the number of state if large enough $\Nstate \gg 1$, the response in equilibrium $\dnds$ due to change in phenotype can be approximated as:

@@

\frac{ \der \dnds\eq}{\der \x\eq}  & = 2 \left[ \frac{\Nstate (\x\eq- 1) + 1+\left[\Nstate (\x\eq-1)^2+2 \x\eq-1\right] \lambda_{K}(\x\eq) }{(\Nstate (\x\eq-1)+1)^2}\right] + \frac{\Nstate - 2}{\Nstate - 1}, \\
& \simeq \frac{2 \lambda_{K}(\x\eq)}{\Nstate} + 1, \\

@@

Define the function $G(\x, \Ne )$ as:

@@

G(\x, \Ne ) \equiv \lambda_K (\x\eq) + 4 \Ne s(\x, \dx), 

@@

The equilibrium equation (eq. \ref{eq:equilibrium}) states that $G(\x\eq, \Ne )=0$, meaning that $\x\eq$ is implicitly a function of $\Ne$:

@@

G(\x\eq(\Ne), \Ne ) = 0, \\
\Longrightarrow \frac{\partial G(\x\eq, \Ne ) }{\partial \x\eq }\frac{ \der \x\eq}{\der \Ne} + \frac{ \partial G(\x\eq, \Ne )}{\partial \Ne} = 0, \\
\iff \left[  \frac{\partial \lambda_K (\x\eq) }{\partial \x\eq }  + 4 \Ne \frac{ \partial s(\x\eq, \dx) }{\partial \x\eq } \right]\frac{ \der \x\eq}{\der \Ne} + 4 s(\x\eq, \dx) = 0, \\
\iff \left[  \frac{\partial \lambda_K (\x\eq) }{\partial \x\eq } + 4 \Ne \frac{ \partial^2 \ln f(\x\eq) }{\partial {\x\eq}^2}\dx \right]\frac{ \der \x\eq}{\der \Ne}  = - 4 \frac{ \partial \ln f(\x\eq) }{\partial {\x\eq}}\dx \text{ from eq. \ref{eq:s_from_fitness}}, \\
\iff 4 \dx \left[ \frac{1}{4 \dx \Ne } \frac{\partial \lambda_K (\x\eq) }{\partial \x\eq } + \frac{ \partial^2 \ln f(\x\eq) }{\partial {\x\eq}^2} \right] \Ne \frac{ \der \x\eq}{\der \Ne}  = - 4 \dx \frac{ \partial \ln f(\x\eq) }{\partial {\x\eq}}, \\
\iff \frac{ \der \x\eq}{\der \ln (\Ne)}  = - \frac{\frac{ \partial \ln f(\x\eq) }{\partial {\x\eq}}}{\frac{1}{4 \dx \Ne} \frac{\partial \lambda_K (\x\eq) }{\partial \x\eq } + \frac{ \partial^2 \ln f(\x\eq) }{\partial {\x\eq}^2}}  \label{eq:dx_dlnNe}.

@@

\frac{\der \dnds\eq}{\der \ln (\Ne)} & = \frac{\der \dnds\eq}{\der \x\eq} \frac{ \der \x\eq}{\der \ln (\Ne)}, \\
 &= - \frac{\der \dnds\eq}{\der \x\eq} \frac{\frac{ \partial \ln f(\x\eq) }{\partial {\x\eq}}}{\frac{1}{4 \dx \Ne } \frac{\partial \lambda_K (\x\eq) }{\partial \x\eq } + \frac{ \partial^2 \ln f(\x\eq) }{\partial {\x\eq}^2}} \text{ from eq. \ref{eq:dx_dlnNe}} \label{eq:dw_dlnNe}.

@@

Moreover, with the approximation that $\left| 4 \Ne \frac{ \partial s(\x\eq, \dx) }{\partial \x\eq } \right| \gg \left| \frac{\partial \lambda_K (\x\eq) }{\partial \x\eq } \right|$, meaning that a change in phenotype causes a higher change in scaled selection coefficient than mutational bias, we have:

@@

\frac{ \der \x\eq}{\der \ln (\Ne)}  = - \frac{\frac{ \partial \ln f(\x\eq) }{\partial {\x\eq}}}{\frac{1}{4 \dx \Ne} \frac{\partial \lambda_K (\x\eq) }{\partial \x\eq } + \frac{ \partial^2 \ln f(\x\eq) }{\partial {\x\eq}^2}},\\
\Longrightarrow \frac{ \der \x\eq}{\der \ln (\Ne)}  \simeq - \frac{\frac{ \partial \ln f(\x\eq) }{\partial {\x\eq}}}{\frac{ \partial^2 \ln f(\x\eq) }{\partial {\x\eq}^2}}. \label{eq:approx_dx_dlnNe}

@@

Together, these approximations leads to the following elasticity in equilibrium $\dnds$ after change in $\Ne$ as:

@@

\frac{ \der \omega\eq}{\der \ln (\Ne)} & \simeq - \frac{\frac{ \partial \ln f(\x\eq) }{\partial {\x\eq}}}{\frac{ \partial^2 \ln f(\x\eq) }{\partial {\x\eq}^2}}\label{eq:approx_dw_dlnNe}

@@

All phenotype-fitness functions considered below are log-concave, and as a result, $\frac{ \partial \ln f(\x\eq) }{\partial {\x\eq}}$ is an decreasing function of $\x$; the less stable the protein already is, the stronger the purifying selection against additional destabilizing mutations. More precisely, fitness functions depends on the folded fraction of the protein of interest, which is given by the Fermi-Dirac distribution:

@@

P_{F} (\x) &=& \frac{1}{1 + e^{\beta(\alpha + \gamma \Nsite \x)}}

@@

where $\x$ is the fraction of destabilizing mutations, each contributing a $\Delta \Delta G = \gamma$, and $\beta = 1 / kT$. Thus, $\alpha = \Delta G_0 < 0$ is the difference in free energy between folded and unfolded state when all sites are stable. $\Nsite \gamma$ is the expected change in $\Delta G$ when all sites are unstable. The misfolded fraction is  $P_{U} = 1-P_{F}$. In addition, $P_{F}$ is typically close to 1 (or $P_{U} \ll 1$), so that we can use a first-order approximation:

@@

P_{F}(\x) &=& 1 - P_{U}(\x) \\
&\simeq& 1 - e^{\beta(\alpha + \gamma \Nsite \x)}

@@

P_{U}(\x) &\simeq& e^{\beta(\alpha + \gamma \Nsite \x)}

@@

f(\x) = \frac{1}{1 + \e^{\beta(\alpha + \Nsite \gamma \x)}}. \label{eq:fitness}

@@

\frac{\partial \ln (f(\x))}{\partial \x}  & = - \frac{\partial \ln \left( 1 + \e^{\beta(\alpha + \Nsite \gamma \x)} \right)}{\partial \x} \text{ from eq. \ref{eq:fitness}}, \\
& = - \beta \Nsite \gamma \frac{\e^{\beta(\alpha + \Nsite \gamma \x)}}{1 + \e^{\beta(\alpha + \Nsite \gamma \x)}} \label{eq:df_dx},\\
& \simeq - \beta \Nsite \gamma \e^{\beta(\alpha + \Nsite \gamma \x)}.

@@

The equilibrium phenotype ($\x\eq$) is :

@@

\lambda_{K}(\x\eq) = 4\Ne \beta \gamma \frac{\e^{\beta(\alpha + \Nsite \gamma \x\eq)}}{1 + \e^{\beta(\alpha + \Nsite \gamma \x\eq)}}  \text{ from eq. \ref{eq:equilibrium_lambda} and \ref{eq:df_dx}}.

@@

\frac{\partial^2 \ln (f(\x))}{\partial \x^2} & = - \beta \Nsite \gamma \frac{\partial}{\partial \x} \left( \frac{\e^{\beta(\alpha + \Nsite \gamma \x)}}{1 + \e^{\beta(\alpha + \Nsite \gamma \x)}} \right) \text{ from eq. \ref{eq:df_dx}}, \\
 & = - \beta \Nsite \gamma  \beta \Nsite \gamma \frac{\e^{\beta(\alpha + \Nsite \gamma \x)}}{\left( 1 + \e^{\beta(\alpha + \Nsite \gamma \x)}\right)^2}, \\
 & = \frac{\beta \Nsite \gamma}{1 + \e^{\beta(\alpha + \Nsite \gamma \x)}} \frac{ \partial \ln f(\x) }{\partial \x} \text{ from eq. \ref{eq:df_dx}} \label{eq:df2_dx2}, \\
 & \simeq \beta \Nsite \gamma \frac{ \partial \ln f(\x) }{\partial \x} 

@@

\frac{ \der \dnds\eq}{\der \ln (\Ne)} & \simeq - \frac{1}{\beta \Nsite \gamma} \text{from eq. \ref{eq:df2_dx2} and \ref{eq:approx_domega_dx}},

@@

which is independent of $\x\eq$, meaning $\omega$ is linearly decreasing with $\Ne$ in log space.

@@

\ln f(\x) = - A y P_{U} (\x)

@@

\frac{\partial \ln (f(\x))}{\partial \x} &\simeq& -A \beta \gamma \Nsite y  e^{\beta(\alpha + \gamma \Nsite \x)}

@@

4 \Ne \frac{\partial \ln (f(\x))}{\partial \x} &\propto& -4 \Ne y e^{\beta(\alpha + \gamma \Nsite \x)}

@@

\frac{ \der \dnds\eq}{\der \ln (\Ne)} & \simeq - \frac{1}{\beta \Nsite \gamma}.

@@

\ln f(\x) &=& - A y \sum_k \pi_k e^{\beta(\alpha + \gamma\Nsite \x + \gamma k)}

@@

&=& - A y  e^{\beta(\alpha + \gamma\Nsite \x)} \sum_k e^{-\rho n} \frac{(\rho n)^k}{k!} e^{\beta \gamma k}

@@

&=& - A y  e^{\beta(\alpha + \gamma\Nsite \x) + \rho \Nsite (e^{\beta \gamma} -1 )}

@@

& \simeq & - A y  e^{\beta(\alpha + \gamma\Nsite \x) + \rho \beta \gamma \Nsite}

@@

4 \Ne \frac{\partial \ln (f(\x))}{\partial \x} &\propto& -4 \Ne y e^{\beta(\alpha + \gamma \Nsite (x + \rho))}
\\ &\propto& -4 \Ne y e^{\beta(\alpha + \gamma\Nsite \x)}

@@

Specifically, the protein is assumed to be regulated so as to reach a level of expression of functional proteins of $y$, and contributes a total benefit $B$ (which depends on its specific function). Given that only a fraction $P_{F}(\x) = 1-P_{U}(\x)$ of the total amount of protein expressed by the cell is functional, the total cost of expression $C$ is then equal to:

@@

C(\x) &=& \frac{y}{P_{F}(\x)}
\\ &\simeq& y (1 + P_{U}(\x))

@@

\ln f (\x) & = & -A \frac{y}{B} \left( 1 + e^{\beta(\alpha + \gamma\Nsite \x)} \right)

@@

&=& -by (1 + e^{\beta(\alpha + \gamma\Nsite \x)})

@@

\ln f(\x) &=& -b \sum_i \epsilon_i

>>>>

For a given genotype sequence, define phenotype $0 \leq x \leq 1$ as the current proportion of sites in the unstable state.
After a mutation, given that only one site can change at a time, the absolute change of $x$ is either $0$ or $\dx=\sfrac{1}{\Nsite}$.
Define $\rho_{x}(\dx)$ as the probability to get a change of phenotype equal to $\dx$, if the current phenotype is $x$:\\

@@

\dx &\text{ with probability } \rho_{x}(\dx) = 1-x, \\
0 &\text{ with probability } \rho_{x}(0) = x \left[1 - \frac{1}{\Nstate - 1}\right], \\
-\dx &\text{ with probability } \rho_{x}(-\dx) = \frac{x}{\Nstate - 1}.\\

@@

$s(x, \dx)$ is the selection coefficient of an effect $\dx$ if the current phenotype is $x$:

@@

s(x, \dx) &= \frac{f(x + \dx) - f(x)}{f( x)}, \\
 &\simeq \frac{1}{f(x)}\frac{ \partial f(x) }{\partial x} \dx, \\
 &\simeq \frac{ \partial \ln f(x) }{\partial x} \dx, \label{eq:s_from_fitness}

@@

where $f( x)$ is the Wrightian fitness of phenotype $x$. And thus we also have:

@@

s(x, -\dx) \simeq - s(x, \dx) \text{ from eq. \ref{eq:s_from_fitness}} \label{eq:s_minus_deltax}, \\
\iff S(x, -\dx) \simeq - S(x, \dx), \label{eq:S_minus_deltax}

@@

where $S(x\eq, \dx) = 4\Ne s (x\eq, \dx)$ is the scaled selection coefficient.

@@

The probability of fixation of a mutation with effect $\dx$, for a resident phenotype $x$ is :

@@

p_{\text{fix}}(x, \dx) &= \frac{ 2 s(x, \dx)}{1 - \e^{-4\Ne s(x, \dx)}}, \\
 &= \frac{ 2 s(x, \dx)}{1 - \e^{-S(x, \dx)}}. \label{eq:pfix}

@@

p_{\text{fix}}(x, 0) = \frac{ 1}{2 \Ne} \label{eq:pfix0}.

@@

\frac{p_{\text{fix}}(x, \dx)}{p_{\text{fix}}(x, 0)} & = \frac{ 2 \Ne 2 s(x, \dx)}{1 - \e^{-S(x, \dx)}} \text{ from eq. \ref{eq:pfix} and \ref{eq:pfix0}},\\
& = \frac{S(x, \dx)}{1 - \e^{-S(x, \dx)}} \label{eq:pfixratio}.

@@

At equilibrium phenotype $x\eq$, the expected selection coefficient of mutation that reached fixation must be $0$:

@@

 0 = \E_{\dx} \left[ s(x\eq, \dx) p_{\text{fix}}(x\eq, \dx) \right], \\
\iff 0 = \frac{ 2 s(x\eq, \dx)^2}{1 - \e^{-S(x\eq, \dx)}}   \rho_{x\eq}(\dx) + s(x\eq, 0) \frac{ \rho_{x\eq}(0)}{2\Ne} + \frac{ 2s(x\eq, -\dx)^2}{1 - \e^{-S(x\eq, -\dx)}} \rho_{x\eq}(-\dx)\text{ from eq. \ref{eq:pfix} and \ref{eq:pfix0}}, \\
\Longrightarrow \frac{ 2s(x\eq, \dx)^2}{1 - \e^{-S(x\eq, \dx)}}   \rho_{x\eq}(\dx) \simeq \frac{ - 2s(x\eq, \dx)^2}{1 - \e^{S(x\eq, \dx)}}   \rho_{x\eq}(-\dx) \text{ from eq. \ref{eq:S_minus_deltax}}, \\
\iff \frac{ \rho_{x\eq}(\dx)}{\rho_{x\eq}(-\dx)} \simeq \e^{-S(x\eq, \dx)} \frac{ \e^{-S(x\eq, \dx)} - 1}{ \e^{-S(x\eq, \dx)} \left( 1 - \e^{S(x\eq, \dx)} \right)}, \\
\iff \ln \left( \frac{1 - x\eq}{x\eq} \right) + \ln (\Nstate-1) \simeq -S(x\eq, \dx) \text{ from eq. \ref{eq:proba}} \label{eq:equilibrium}, \\
\iff \lambda_{K}(x\eq) \simeq -S(x\eq, \dx) \label{eq:equilibrium_lambda},

@@

where  $\lambda_{K}(x\eq) = \ln \left( \frac{1 - x\eq}{x\eq} \right) + \ln (\Nstate-1)$.
\subsection{Substitution rate bias ($\bm{\omega}$) at equilibrium}
The substitution rate of selected mutations relative to neutral mutations ($\omega$) is by definition:

@@

\omega & = \E_{\dx} \left[ \frac{p_{\text{fix}}(x, \dx)}{p_{\text{fix}}(x, 0)} \right], \\
 & = (1 - x) \frac{ S(x, \dx)}{1 - \e^{-S(x, \dx)}} + x \left(\frac{\Nstate - 2}{\Nstate - 1}\right) + \frac{x}{\Nstate-1} \frac{ S(x, -\dx)}{1 - \e^{-S(x, -\dx)}} \text{ from eq. \ref{eq:proba}, \ref{eq:pfix} and \ref{eq:pfix0}}, \\
 & = (1 - x) \frac{ S(x, \dx)}{1 - \e^{-S(x, \dx)}} - \frac{x}{\Nstate-1}  \frac{ S(x, \dx)}{1 - \e^{S(x, \dx)}} +  x \left(\frac{\Nstate - 2}{\Nstate - 1}\right) \text{ from eq. \ref{eq:S_minus_deltax}}.

@@

$\omega\eq$ at equilibrium is then determined by the phenotype at equilibrium $x\eq$:

@@

\omega\eq &= (1 - x\eq) \frac{ S(x\eq, \dx)}{1 - \e^{-S(x\eq, \dx)}} - \frac{x\eq}{\Nstate-1} \frac{ S(x\eq, \dx)}{1 - \e^{S(x\eq, \dx)}} + x\eq \left(\frac{\Nstate - 2}{\Nstate - 1}\right) , \\
 &= x\eq \left[ \frac{2 (x\eq-1)  \lambda_{K}(x\eq) }{\Nstate (x\eq-1)+1} + \frac{\Nstate - 2}{\Nstate - 1}\right] \text{ from eq. \ref{eq:equilibrium}}.

@@

Moreover, given that the number of state if large enough $\Nstate \gg 1$, the equilibrium $\omega$ can be approximated as:

@@

\omega\eq  & = x\eq \left[ \frac{2 (x\eq-1)  \lambda_{K}(x\eq) }{\Nstate (x\eq-1)+1} + \frac{\Nstate - 2}{\Nstate - 1}\right], \\
& \simeq x\eq

@@

And the derivative of $\omega\eq$ w.r.t to $x\eq$ is: 

@@

\frac{\der \omega\eq}{\der x\eq} = 2 \left[ \frac{\Nstate (x\eq- 1) + 1+\left[\Nstate (x\eq-1)^2+2 x\eq-1\right] \lambda_{K}(x\eq) }{(\Nstate (x\eq-1)+1)^2}\right] + \frac{\Nstate - 2}{\Nstate - 1} \label{eq:dw_dx}.

@@

Moreover, given that the number of state if large enough $\Nstate \gg 1$, the response in equilibrium $\omega$ due to change in phenotype can be approximated as:

@@

\frac{ \der \omega\eq}{\der x\eq}  & = 2 \left[ \frac{\Nstate (x\eq- 1) + 1+\left[\Nstate (x\eq-1)^2+2 x\eq-1\right] \lambda_{K}(x\eq) }{(\Nstate (x\eq-1)+1)^2}\right] + \frac{\Nstate - 2}{\Nstate - 1}, \\
& \simeq \frac{2 \lambda_{K}(x\eq)}{\Nstate} + 1, \\

@@

Define the function $G(x, \Ne )$ as:

@@

G(x, \Ne ) \equiv \lambda_K (x\eq) + 4 \Ne s(x, \dx), 

@@

The equilibrium equation (eq. \ref{eq:equilibrium}) states that $G(x\eq, \Ne )=0$, meaning that $x\eq$ is implicitly a function of $\Ne$:

@@

G(x\eq(\Ne), \Ne ) = 0, \\
\Longrightarrow \frac{\partial G(x\eq, \Ne ) }{\partial x\eq }\frac{ \der x\eq}{\der \Ne} + \frac{ \partial G(x\eq, \Ne )}{\partial \Ne} = 0, \\
\iff \left[  \frac{\partial \lambda_K (x\eq) }{\partial x\eq }  + 4 \Ne \frac{ \partial s(x\eq, \dx) }{\partial x\eq } \right]\frac{ \der x\eq}{\der \Ne} + 4 s(x\eq, \dx) = 0, \\
\iff \left[  \frac{\partial \lambda_K (x\eq) }{\partial x\eq } + 4 \Ne \frac{ \partial^2 \ln f(x\eq) }{\partial {x\eq}^2}\dx \right]\frac{ \der x\eq}{\der \Ne}  = - 4 \frac{ \partial \ln f(x\eq) }{\partial {x\eq}}\dx \text{ from eq. \ref{eq:s_from_fitness}}, \\
\iff 4 \dx \left[ \frac{1}{4 \dx \Ne } \frac{\partial \lambda_K (x\eq) }{\partial x\eq } + \frac{ \partial^2 \ln f(x\eq) }{\partial {x\eq}^2} \right] \Ne \frac{ \der x\eq}{\der \Ne}  = - 4 \dx \frac{ \partial \ln f(x\eq) }{\partial {x\eq}}, \\
\iff \frac{ \der x\eq}{\der \ln (\Ne)}  = - \frac{\frac{ \partial \ln f(x\eq) }{\partial {x\eq}}}{\frac{1}{4 \dx \Ne} \frac{\partial \lambda_K (x\eq) }{\partial x\eq } + \frac{ \partial^2 \ln f(x\eq) }{\partial {x\eq}^2}}  \label{eq:dx_dlnNe}.

@@

\frac{\der \omega\eq}{\der \ln (\Ne)} & = \frac{\der \omega\eq}{\der x\eq} \frac{ \der x\eq}{\der \ln (\Ne)}, \\
 &= - \frac{\der \omega\eq}{\der x\eq} \frac{\frac{ \partial \ln f(x\eq) }{\partial {x\eq}}}{\frac{1}{4 \dx \Ne } \frac{\partial \lambda_K (x\eq) }{\partial x\eq } + \frac{ \partial^2 \ln f(x\eq) }{\partial {x\eq}^2}} \text{ from eq. \ref{eq:dx_dlnNe}} \label{eq:dw_dlnNe}.

@@

Moreover, with the approximation that $\left| 4 \Ne \frac{ \partial s(x\eq, \dx) }{\partial x\eq } \right| \gg \left| \frac{\partial \lambda_K (x\eq) }{\partial x\eq } \right|$, meaning that a change in phenotype causes a higher change in scaled selection coefficient than mutational bias, we have:

@@

\frac{ \der x\eq}{\der \ln (\Ne)}  = - \frac{\frac{ \partial \ln f(x\eq) }{\partial {x\eq}}}{\frac{1}{4 \dx \Ne} \frac{\partial \lambda_K (x\eq) }{\partial x\eq } + \frac{ \partial^2 \ln f(x\eq) }{\partial {x\eq}^2}},\\
\Longrightarrow \frac{ \der x\eq}{\der \ln (\Ne)}  \simeq - \frac{\frac{ \partial \ln f(x\eq) }{\partial {x\eq}}}{\frac{ \partial^2 \ln f(x\eq) }{\partial {x\eq}^2}}. \label{eq:approx_dx_dlnNe}

@@

Together, these approximations leads to the following elasticity in equilibrium $\omega$ after change in $\Ne$ as:

@@

\frac{ \der \omega\eq}{\der \ln (\Ne)} & \simeq - \frac{\frac{ \partial \ln f(x\eq) }{\partial {x\eq}}}{\frac{ \partial^2 \ln f(x\eq) }{\partial {x\eq}^2}}\label{eq:approx_dw_dlnNe}

@@

All phenotype-fitness functions considered below are log-concave, and as a result, $\frac{ \partial \ln f(x\eq) }{\partial {x\eq}}$ is an decreasing function of $x$; the less stable the protein already is, the stronger the purifying selection against additional destabilizing mutations. More precisely, fitness functions depends on the folded fraction of the protein of interest, which is given by the Fermi-Dirac distribution:

@@

P_{F} (x) &=& \frac{1}{1 + e^{\beta(\alpha + \gamma \Nsite x)}}

@@

where $x$ is the fraction of destabilizing mutations, each contributing a $\Delta \Delta G = \gamma$, and $\beta = 1 / kT$. Thus, $\alpha = \Delta G_0 < 0$ is the difference in free energy between folded and unfolded state when all sites are stable. $\Nsite \gamma$ is the expected change in $\Delta G$ when all sites are unstable. The misfolded fraction is  $P_{U} = 1-P_{F}$. In addition, $P_{F}$ is typically close to 1 (or $P_{U} \ll 1$), so that we can use a first-order approximation:

@@

P_{F}(x) &=& 1 - P_{U}(x) \\
&\simeq& 1 - e^{\beta(\alpha + \gamma \Nsite x)}

@@

P_{U}(x) &\simeq& e^{\beta(\alpha + \gamma \Nsite x)}

@@

f(x) = \frac{1}{1 + \e^{\beta(\alpha + \Nsite \gamma x)}}. \label{eq:fitness}

@@

\frac{\partial \ln (f(x))}{\partial x}  & = - \frac{\partial \ln \left( 1 + \e^{\beta(\alpha + \Nsite \gamma x)} \right)}{\partial x} \text{ from eq. \ref{eq:fitness}}, \\
& = - \beta \Nsite \gamma \frac{\e^{\beta(\alpha + \Nsite \gamma x)}}{1 + \e^{\beta(\alpha + \Nsite \gamma x)}} \label{eq:df_dx},\\
& \simeq - \beta \Nsite \gamma \e^{\beta(\alpha + \Nsite \gamma x)}.

@@

The equilibrium phenotype ($x\eq$) is :

@@

\lambda_{K}(x\eq) = 4\Ne \beta \gamma \frac{\e^{\beta(\alpha + \Nsite \gamma x\eq)}}{1 + \e^{\beta(\alpha + \Nsite \gamma x\eq)}}  \text{ from eq. \ref{eq:equilibrium_lambda} and \ref{eq:df_dx}}.

@@

\frac{\partial^2 \ln (f(x))}{\partial x^2} & = - \beta \Nsite \gamma \frac{\partial}{\partial x} \left( \frac{\e^{\beta(\alpha + \Nsite \gamma x)}}{1 + \e^{\beta(\alpha + \Nsite \gamma x)}} \right) \text{ from eq. \ref{eq:df_dx}}, \\
 & = - \beta \Nsite \gamma  \beta \Nsite \gamma \frac{\e^{\beta(\alpha + \Nsite \gamma x)}}{\left( 1 + \e^{\beta(\alpha + \Nsite \gamma x)}\right)^2}, \\
 & = \frac{\beta \Nsite \gamma}{1 + \e^{\beta(\alpha + \Nsite \gamma x)}} \frac{ \partial \ln f(x) }{\partial x} \text{ from eq. \ref{eq:df_dx}} \label{eq:df2_dx2}, \\
 & \simeq \beta \Nsite \gamma \frac{ \partial \ln f(x) }{\partial x} 

@@

\frac{ \der \omega\eq}{\der \ln (\Ne)} & \simeq - \frac{1}{\beta \Nsite \gamma} \text{from eq. \ref{eq:df2_dx2} and \ref{eq:approx_domega_dx}},

@@

which is independent of $x\eq$, meaning $\omega$ is linearly decreasing with $\Ne$ in log space.

@@

\ln f(x) = - A y P_{U} (x)

@@

\frac{\partial \ln (f(x))}{\partial x} &\simeq& -A \beta \gamma \Nsite y  e^{\beta(\alpha + \gamma \Nsite x)}

@@

4 \Ne \frac{\partial \ln (f(x))}{\partial x} &\propto& -4 \Ne y e^{\beta(\alpha + \gamma \Nsite x)}

@@

\frac{ \der \omega\eq}{\der \ln (\Ne)} & \simeq - \frac{1}{\beta \Nsite \gamma}.

@@

\ln f(x) &=& - A y \sum_k \pi_k e^{\beta(\alpha + \gamma\Nsite x + \gamma k)}

@@

&=& - A y  e^{\beta(\alpha + \gamma\Nsite x)} \sum_k e^{-\rho n} \frac{(\rho n)^k}{k!} e^{\beta \gamma k}

@@

&=& - A y  e^{\beta(\alpha + \gamma\Nsite x) + \rho \Nsite (e^{\beta \gamma} -1 )}

@@

& \simeq & - A y  e^{\beta(\alpha + \gamma\Nsite x) + \rho \beta \gamma \Nsite}

@@

4 \Ne \frac{\partial \ln (f(x))}{\partial x} &\propto& -4 \Ne y e^{\beta(\alpha + \gamma \Nsite (x + \rho))}
\\ &\propto& -4 \Ne y e^{\beta(\alpha + \gamma\Nsite x)}

@@

Specifically, the protein is assumed to be regulated so as to reach a level of expression of functional proteins of $y$, and contributes a total benefit $B$ (which depends on its specific function). Given that only a fraction $P_{F}(x) = 1-P_{U}(x)$ of the total amount of protein expressed by the cell is functional, the total cost of expression $C$ is then equal to:

@@

C(x) &=& \frac{y}{P_{F}(x)}
\\ &\simeq& y (1 + P_{U}(x))

@@

\ln f (x) & = & -A \frac{y}{B} \left( 1 + e^{\beta(\alpha + \gamma\Nsite x)} \right)

@@

&=& -by (1 + e^{\beta(\alpha + \gamma\Nsite x)})

@@

\ln f(x) &=& -b \sum_i \epsilon_i
